---
title: "Reproducible Research: Peer Assignment 1"
output:
  html_document:
    keep_md: true
---

## Load and Process Data
```{r}
setwd("D:/Coursera")
stepData <- read.csv("activity.csv")
stepData$date <- as.Date(stepData$date, format = "%Y-%m-%d")
```

## What is the total number of steps taken per day?
rm.na = TRUE to ignore missing values.
```{r}
steps <- aggregate(steps ~ date, rm.na = TRUE, data = stepData, FUN = sum)
```

### Histogram of steps per day
```{r}
plot(steps, type = "h", lwd = 5,
	xlab = "Date", ylab = "Steps", main = "Total Steps per Day")
```

### Calculate and report the mean and median of the total number of steps per day
```{r}
aggregate(steps ~ date, data = stepData, FUN = mean)
aggregate(steps ~ date, data = stepData, FUN = median)
```
Note that the median function does not seem to be working properly with data tables.  It returns all 0's here.

## What is the average daily activity pattern?
```{r}
plot(aggregate(steps ~ interval, data = stepData, FUN = mean), type = "l",
	xlab = "Interval", ylab = "Steps", main = "Average Number of Steps Across All Days")
```

### Which 5-minute interval, on average across all days, contains the maximum number of steps?
```{r}
max(stepData$steps, na.rm = TRUE)
```

## Imputing Missing Values
### Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(stepData))
```

### Fill in missing data
I'm using the mean steps of the original dataset as a constant to fill in every missing value.
```{r}
stepData <- read.csv("activity.csv")
stepData2 <- stepData
meanStep <- mean(na.omit(stepData$steps))
stepData2$steps[is.na(stepData2$steps)] <- meanStep
stepData2$date <- as.Date(stepData2$date, format = "%Y-%m-%d")
```

### Make a histogram of total number of steps for each day with filled dataset
```{r}
steps2 <- aggregate(steps ~ date, data = stepData2, FUN = sum)

par(mfrow = c(1, 2))
plot(steps, type = "h", lwd = 5, xlab = "Date", ylab = "Steps",
	main = "Total Steps with NAs")
plot(steps2, type = "h", lwd = 5, xlab = "Date", ylab = "Steps",
	main = "Total Steps with NAs Filled")
```
The filled dataset has less gaps overall, making for a somewhat smoother curve.

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
Using the filled dataset
```{r}
stepData2$dayOfWeek <- factor(format(stepData2$date, "%A"))
levels(stepData2$dayOfWeek) <- list(weekday = c("Monday", "Tuesday", "Wednesday",
											                        "Thursday, Friday"),
								                    weekend = c("Saturday", "Sunday"))
```

### Make a times series panel plot of average number of steps taken across all weekdays or weekend days.
```{r}
par(mfrow = c(2, 1))
with(stepData2[stepData2$dayOfWeek == "weekday",],
	plot(aggregate(steps ~ interval, FUN = mean), type = "l",
		xlab = "Interval", ylab = "Steps", main = "Weekday Step Totals"))
with(stepData2[stepData2$dayOfWeek == "weekend",],
	plot(aggregate(steps ~ interval, FUN = mean), type = "l",
		xlab = "Interval", ylab = "Steps", main = "Weekend Step Totals"))
```